---
title: "newdraft"
output: html_document
date: "2025-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)        # For reading CSV files
library(dplyr)        # Data manipulation
library(tidyr)        # Data tidying and handling missing values
library(tidyverse)    # Collection of R packages including ggplot2, dplyr, tidyr
library(caret)        # For data partitioning and machine learning models
library(glmnet)       # Regularization methods (Ridge, Lasso)
library(ISLR)         # Contains datasets and modeling functions
library(ggplot2)      # For data visualization (optional but useful)
library(stringr)      # For string manipulation (gsub alternatives)
library(forcats)      # For working with categorical variables (factors)
library(purrr)    
library(randomForest) # Functional programming tools (used in lapply-like functions)
library(xgboost)
library(nnet)
library(fastDummies)
library(e1071)
library(FNN) 
```


## Cleaning the Data

```{r message=FALSE, warning=FALSE}

all_car_adverts <- read_csv("all_car_adverts.csv", 
col_types = cols(car_price = col_number(), 
car_seller_rating = col_number(), miles = col_number(), year= col_number()))

# Convert PS to BHP where needed (only modify engine_size when the unit is "ps")
all_car_adverts$engine_size <- ifelse(
  all_car_adverts$engine_size_unit == "ps", 
  all_car_adverts$engine_size * 0.98632,  # Convert only PS values
  all_car_adverts$engine_size  # Keep BHP values unchanged
)
all_car_adverts$engine_size_unit[all_car_adverts$engine_size_unit == "ps"] <- "bhp"
all_car_adverts$reg <- gsub(" reg", "", all_car_adverts$reg)

all_car_adverts <- all_car_adverts %>%
  mutate(luxury_level = case_when(
    # Super Luxury (5)
    make %in% c("Aston Martin", "Bentley", "Ferrari", "Lamborghini", "Maserati", "McLaren", "Rolls-Royce") ~ 5,
    
    # Luxury (4)
    make %in% c("Alfa Romeo", "Audi", "BMW", "Cadillac", "Jaguar", "Land Rover", "Lexus", "Mercedes-Benz", "Porsche", "Volvo") ~ 4,
    
    # Upper Mid-range (3)
    make %in% c("Abarth", "Chrysler", "Cupra", "Dodge", "DS AUTOMOBILES", "Infiniti", "Lotus", "MINI", "Morgan", "Peugeot", "Volkswagen") ~ 3,
    
    # Mid-range (2)
    make %in% c("Chevrolet", "Citroen", "Fiat", "Ford", "Honda", "Hummer", "Hyundai", "Jeep", "Kia", "Mazda", "Nissan", "Renault", "SEAT", "SKODA", "Subaru", "Toyota", "Vauxhall") ~ 2,
    
    # Economic (1)
    make %in% c("Aixam", "Austin", "Caterham", "Dacia", "Daewoo", "Daihatsu", "Daimler", "Isuzu", "Lancia", "London Taxis International", "MG", "Mitsubishi", "Perodua", "Proton", "Rover", "Saab", "Smart", "Ssangyong", "Suzuki", "TVR") ~ 1
  ))



all_car_adverts <- all_car_adverts %>% select(-c(car_badges, car_specs, brand_new, discounted, car_attention_grabber,car_sub_title, car_seller, car_seller_location,...1, engine_size_unit, car_title, reg))
all_car_adverts <- na.omit(all_car_adverts)
all_car_adverts <- all_car_adverts %>%
  rename(fuel_type = feul_type)

########full_service, full_dealership part_service, part_warranty, first_year_road_tax, first_year_road_tax and ulez I included. Ulez is ultra low emissions zone, and electric cars comply, but not all petrol cars. Might be worth keeping this one. But of course you can always delete these columns again for your analysis! But for the first data partition I wanted to include them so that I can use them for my models as well. ############

### I also threw out reg as we never use it


# Convert categorical variables to factors
factor_cols <- names(Filter(is.character, all_car_adverts))
all_car_adverts[factor_cols] <- lapply(all_car_adverts[factor_cols], as.factor)

# Handle zero values in `car_price` to prevent log(0) issues
all_car_adverts <- all_car_adverts %>% filter(car_price > 0)

# Apply log transformation to car_price
all_car_adverts <- all_car_adverts %>% mutate(log_price = log(car_price))
```

## Classification using AI:
# Prompt:
Classify the following car brands into five luxury levels from 1 (Economic) to 5 (Super Luxury).

The classification should always be consistent with the given reference.
Use the following criteria:
5 - Super Luxury: Exclusive luxury and exotic brands (ultra-high price, exclusivity).
4 - Luxury: Premium brands offering high-end luxury models.
3 - Upper Mid-range: Brands with some luxury or premium models.
2 - Mid-range: Popular mainstream brands with mass-market appeal.
1 - Economic: Entry-level or budget-friendly manufacturers.

Instructions:

If a brand is not listed, classify it by similarity to listed brands, but prioritize sticking to the above reference.
Output the classification as a table with two columns: Brand and Luxury Level (1-5).
Ensure reproducibility by not altering the categories in future runs.

Car Brands: 
Aston Martin, Bentley, Ferrari, Lamborghini, Maserati, McLaren, Rolls-Royce
Alfa Romeo, Audi, BMW, Cadillac, Jaguar, Land Rover, Lexus, Mercedes-Benz, Porsche, Volvo
Abarth, Chrysler, Cupra, Dodge, DS AUTOMOBILES, Infiniti, Lotus, MINI, Morgan, Peugeot, Volkswagen
Chevrolet, Citroen, Fiat, Ford, Honda, Hummer, Hyundai, Jeep, Kia, Mazda, Nissan, Renault, SEAT, SKODA, Subaru, Toyota, Vauxhall
Aixam, Austin, Caterham, Dacia, Daewoo, Daihatsu, Daimler, Isuzu, Lancia, London Taxis International, MG, Mitsubishi, Perodua, Proton, Rover, Saab, Smart, Ssangyong, Suzuki, TVR


```{r}
unique_brands_luxury <- all_car_adverts %>%
  select(make, luxury_level) %>%    # Select only relevant columns
  distinct() %>%                    # Get unique combinations
  arrange(desc(luxury_level), make) # Sort by luxury level (descending) and brand name

# View the result
unique_brands_luxury

```
## Setting up the first data partition for a holy test set and the rest, with which we will train different models with training and validation sets. 

```{r}
set.seed(5462)


#So first I set aside 10% of the data as a holy test set, only to be used at the end of creating each model. With the remaining 85%, we can each play around with how much of that 90% we use as training or testing data, so we create our own sample divisions with that 90%. And also we might each select or use different columns. But the 10% is only for the end

#I also sampled randomly, not stratified with luxury level. My thinking was that we could first sample randomly, and then later try different variables to stratify with as robustness checks



test_index <- sample(seq_len(nrow(all_car_adverts)), size = 0.03 * nrow(all_car_adverts))
final_test_set <- all_car_adverts[test_index, ]
remaining_data <- all_car_adverts[-test_index, ]  # 85% left for training/validation

#To make sure that all test sets have the same observattions as for the randdom forest, we will delete firstly the 



```



Random forest model - final


```{r}

###### FInal randdom forest mode;l. This one takes a long time to calculate 

# Prepare the training data
forestdata <- remaining_data %>%
   mutate(make = fct_lump(make, n = 52)) %>%  # Keep the 52 most frequent brands
   sample_n(30000) %>%  # Take a random sample of 30,000 rows
   mutate(across(where(is.character), as.factor)) %>%  # Convert character columns to factors
   select(-log_price, -model, -variant, -luxury_level)  


# Prepare the final test set, only include brands that are also in training data
final_test_setRF <- final_test_set %>%
   filter(make %in% forestdata$make) %>%##SO we have the same test set as other mettthods
   mutate(make = factor(make, levels = levels(forestdata$make))) %>%  # Drop unused levels
   select(-log_price, -model, -variant, -luxury_level)

# Define mtry values to test
mtry_valuesRF <- c(3, 6, 9, 12)

# Store models and results
rf_models <- list()
resultsRF <- data.frame(mtry = numeric(), MSE = numeric(), R2 = numeric())

# Loop through different mtry values
for (m in mtry_valuesRF) {
  # Train Random Forest model
  rf_model <- randomForest(
    car_price ~ .,  
    data = forestdata,  
    mtry = m,  
    ntree = 300  # Number of trees
  )

  # Store the trained model
  rf_models[[paste0("mtry_", m)]] <- rf_model

  # Predict on final test set
  rf_preds <- predict(rf_model, newdata = final_test_setRF)

  # Calculate MSE and R²
  mse <- mean((final_test_setRF$car_price - rf_preds)^2)
  r2 <- cor(final_test_setRF$car_price, rf_preds)^2

  # Store results
  resultsRF <- rbind(resultsRF, data.frame(mtry = m, MSE = mse, R2 = r2))
}

# Print results
print(resultsRF)

```



Because the final test set with the random forest excluded two observations, for consistency we will exclude those from the other 
test sets as well. Therefore, we identify the rows that have been deleted and mark them, based on their unique combination of price and miles.  

```{r}
final_test_setex <- final_test_set %>%
   select(-log_price, -model, -variant, -luxury_level)


removeprice <- setdiff(final_test_setex, final_test_setRF)$car_price
removemiles <- setdiff(final_test_setex, final_test_setRF)$miles


#The following line of dplyr code will be applied to the preparation of each test set:

# filter(!(car_price %in% removeprice & miles %in% removemiles))

```


```{r}


#Random forest using out of bag error

rf_modeloob <- randomForest(
  car_price ~ .,        # Formula: Predict car_price using all other variables
  data = forestdata,   # Training dataset
  ntree = 300,          # Number of trees
  mtry = floor(sqrt(ncol(forestdata) - 1)),  # Number of predictors randomly selected at each split
  importance = TRUE,    # Compute variable importance
  keep.forest = TRUE,   # Keep the forest for predictions
  keep.inbag = TRUE     # Keep track of OOB samples
)

# Print the model to view the OOB error rate
print(rf_modeloob)

# Extract and plot the OOB error
plot(rf_modeloob, main = "OOB Error vs Number of Trees")
```

As the MSE does not decrease much after 6, we will use 6. 

```{r}

# Extract the model with mtry = 6
rf_model_6 <- rf_models[["mtry_6"]]

# Plot variable importance
varImpPlot(rf_model_6, main = "Variable Importance (mtry = 6)")

# Plot error rate
plot(rf_model_6, main = "Random Forest Error Rate (mtry = 6)")


```

```{r}

# Predict on final test set
rf_preds_6 <- predict(rf_modeloob, newdata = final_test_setRF)

# Create a dataframe to compare actual vs. predicted values
comparison_RF <- data.frame(
  Actual = final_test_setRF$car_price,
  Predicted = rf_preds_6
)


# Plot Actual vs. Predicted values
library(ggplot2)
ggplot(comparison_RF, aes(x = Actual, y = Predicted)) +
  geom_point(alpha = 0.5) +  # Scatter plot
  geom_abline(slope = 1, intercept = 0, color = "red") +  # Ideal line (y = x)
  theme_minimal() +
  labs(title = "Actual vs Predicted Car Prices using a random forest",
       x = "Actual Price",
       y = "Predicted Price")

```





Neural network model - Final 




```{r}

#Cleaner way 

nndatatrain <- remaining_data %>%
  select(-model, -variant, -log_price, -luxury_level) %>%  # Remove unwanted columns first #Also luxury level
  dummy_cols(select_columns = names(select(., where(is.factor))), 
             remove_first_dummy = TRUE, 
             remove_selected_columns = TRUE) %>%
  mutate(across(where(is.numeric) & !where(~ all(. %in% c(0, 1))), scale))


nndatatest <- final_test_set %>%
  filter(!(car_price %in% removeprice & miles %in% removemiles)) %>% #Remove missing rows from other methods
  select(-model, -variant, -log_price, -luxury_level) %>%  # Remove unwanted columns first #Also luxury level
  dummy_cols(select_columns = names(select(., where(is.factor))), 
             remove_first_dummy = TRUE, 
             remove_selected_columns = TRUE) %>%
  mutate(across(where(is.numeric) & !where(~ all(. %in% c(0, 1))), scale)) 



```


See which decay is best for neural network:

```{r}
train_control <- trainControl(method = "cv", number = 5)  # 5-fold CV

# Train the model while tuning decay
nn_tuned <- train(
  car_price ~ ., 
  data = nndatatrain,  
  method = "nnet",
  trControl = train_control,  
  tuneGrid = expand.grid(size = c(3, 6, 9), decay = c(0, 0.0001, 0.001, 0.01, 0.1)),
  linout = TRUE,
  maxit = 500
)

# View results
print(nn_tuned)
plot(nn_tuned)  # Visualize the effect of decay
```





```{r}

#This takes a while to run 

cv_control <- trainControl(method = "cv", number = 5)  

# Train the neural network model using cross-validation
nn_model_cv <- train(
  car_price ~ ., 
  data = nndatatrain, 
  method = "nnet",
  trControl = cv_control,
  tuneGrid = expand.grid(size = 6, decay = 0.01),  # Hyperparameter tuning
  maxit = 800,  # Number of iterations
  linout = TRUE
)

```

```{r}

# Reverse scaling for predicted prices  
predicted_pricesNN <- predict(nn_model, nndatatest) %>% 
  as.vector() * sd(all_car_adverts$car_price, na.rm = TRUE) + mean(all_car_adverts$car_price, na.rm = TRUE)

# Reverse scaling for actual prices  
actual_pricesNN <- nndatatest$car_price * sd(all_car_adverts$car_price, na.rm = TRUE) + 
                   mean(all_car_adverts$car_price, na.rm = TRUE)

# Compute MSE  
mse_nn <- mean((actual_pricesNN - predicted_pricesNN)^2)

# Print MSE  
print(mse_nn)


```

next try xgboost, svr and knn 

```{r}

#For Extreme gradient boosting, we also need to transform factors to dummies and scale 

xgbdatatrain <- remaining_data %>%
  select(-model, -variant, -log_price, -luxury_level) %>%  # Remove unwanted columns first #Also luxury level
  dummy_cols(select_columns = names(select(., where(is.factor))), 
             remove_first_dummy = TRUE, 
             remove_selected_columns = TRUE) 
 


xgbdatatest <- final_test_set %>%
  filter(!(car_price %in% removeprice & miles %in% removemiles)) %>% #remove rows that arent in others
  select(-model, -variant, -log_price, -luxury_level) %>%  # Remove unwanted columns first #Also luxury level
  dummy_cols(select_columns = names(select(., where(is.factor))), 
             remove_first_dummy = TRUE, 
             remove_selected_columns = TRUE) 

  
```

```{r}

#Now that are training and test sets are done, we can start preparing them for XGBoost

# Convert training and test sets to XGBoost DMatrix

xgbdatatrainmatrix <- xgbdatatrain %>%
  select(-car_price) %>%
  as.matrix() %>%
  xgb.DMatrix(label = xgbdatatrain$car_price)

xgbdatatestmatrix <- xgbdatatest %>%
  select(-car_price) %>%
  as.matrix() %>%
  xgb.DMatrix(label = xgbdatatest$car_price)

```

```{r}
#Now we can train the model

#This chunk takes a few minutes to run

xgb_model <- xgboost(
  data = xgbdatatrainmatrix, 
  nrounds = 100, 
  objective = "reg:squarederror", 
  verbose = 0  # Suppress training output
)



#Instead use cross validation for the xgboost

train_control <- trainControl(method = "cv", number = 5)  

# Train XGBoost with CV
xgb_model <- train(
  car_price ~ ., 
  data = xgbdatatrainmatrix,  
  method = "xgbTree",  
  trControl = train_control
)


```

```{r}
# Define XGBoost parameters
params <- list(
  objective = "reg:squarederror",  # Use squared error for regression
  eval_metric = "rmse",            # Root Mean Squared Error as evaluation metric
  max_depth = 6,                    # Depth of trees
  eta = 0.1                         # Learning rate
)

# Perform 5-fold Cross-Validation
cv_results <- xgb.cv(
  params = params,
  data = xgbdatatrainmatrix,  # Use the pre-processed training matrix
  nrounds = 600,              # Maximum boosting rounds
  nfold = 5,                  # 5-fold cross-validation
  stratified = FALSE,         # Not needed for regression
  verbose = TRUE,
  early_stopping_rounds = 10  # Stop if no improvement after 10 rounds
)

# Print the best number of iterations
print(cv_results$best_iteration)
```
```{r}

#ISolate the best xgb 

best_nrounds <- cv_results$best_iteration

xgb_model <- xgb.train(
  params = params,
  data = xgbdatatrainmatrix,
  nrounds = 300 #Or best iteration 
)
```



```{r}

#And predict the results 

predxgb <- predict(xgb_model, xgbdatatestmatrix)

# Calculate Test MSE

predxgb <- as.numeric(predxgb)  # Ensure predictions are numeric
xgbmse <- mean((xgbdatatest$car_price - predxgb)^2)
cat("Test MSE with XGB:", xgbmse, "\n")

```




Now for SVR


```{r}
library(fastDummies)

# Step 1: Preprocess Training Data
svrdatatrain <- remaining_data %>%
  sample_n(80000) %>%
  select(-model, -variant, -log_price, -luxury_level) %>%  # Remove unwanted columns
  dummy_cols(select_columns = names(select(., where(is.factor))), 
             remove_first_dummy = TRUE, 
             remove_selected_columns = TRUE) %>%
  mutate(across(where(is.numeric) & !where(~ all(. %in% c(0, 1))), scale))

# Step 2: Preprocess Test Data (Ensure Same Structure)
svrdatatest <- final_test_set %>%
  filter(make %in% unique(remaining_data$make)) %>%  # Keep only brands in training
  mutate(make = factor(make, levels = levels(remaining_data$make))) %>%
  select(-model, -variant, -log_price, -luxury_level) %>%
  dummy_cols(select_columns = names(select(., where(is.factor))), 
             remove_first_dummy = TRUE, 
             remove_selected_columns = TRUE) %>%
  mutate(across(where(is.numeric) & !where(~ all(. %in% c(0, 1))), scale))

# Step 3: Match Test Columns to Training Data
missing_cols <- setdiff(names(svrdatatrain), names(svrdatatest))
svrdatatest[missing_cols] <- 0  # Add missing columns as zeros
svrdatatest <- svrdatatest[, names(svrdatatrain)]  # Reorder to match training data

# Step 4: Train SVR Model
library(e1071)
svr_model <- svm(car_price ~ ., data = svrdatatrain, type = "eps-regression", kernel = "radial")

# Step 5: Make Predictions
svr_predictions <- predict(svr_model, newdata = svrdatatest)
actual_prices <- svrdatatest$car_price * price_sd + price_mean

# Calculate MSE and RMSE
svrmse <- mean((svr_predictions - actual_prices)^2)
svrrmse <- sqrt(svrmse)

# Print results
cat("MSE:", svrmse, "\nRMSE:", svrrmse, "\n")

```


```{r}
train_control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

# Train the SVM model using the specified cross-validation
svr_model_cv <- train(
  car_price ~ ., 
  data = svrdatatrain, 
  method = "svmRadial",  # Specify that we're using radial basis kernel
  trControl = train_control,  # Cross-validation settings
  tuneLength = 5  # Number of tuning parameters to try (optional)
)

# View the results
print(svr_model_cv)
```






Now for KNN


```{r}

knndatatrain <- remaining_data %>%
  select(-model, -variant, -log_price, -luxury_level) %>%  # Remove unwanted columns first #Also luxury level
  dummy_cols(select_columns = names(select(., where(is.factor))), 
             remove_first_dummy = TRUE, 
             remove_selected_columns = TRUE) %>%
  mutate(across(where(is.numeric) & !where(~ all(. %in% c(0, 1))), scale))

knndatatest <- final_test_set %>%
  filter(!(car_price %in% removeprice & miles %in% removemiles)) %>%  #remove 
  select(-model, -variant, -log_price, -luxury_level) %>%  # Remove unwanted columns first #Also luxury level
  dummy_cols(select_columns = names(select(., where(is.factor))), 
             remove_first_dummy = TRUE, 
             remove_selected_columns = TRUE) %>%
  mutate(across(where(is.numeric) & !where(~ all(. %in% c(0, 1))), scale)) %>%
  filter(!car_price %in% c(4495, 6995)) #exclude the columns needed 

```


```{r} 

# Extract predictor variables and target variable
knntrain_X <- knndatatrain %>% select(-car_price)
knntrain_Y <- knndatatrain$car_price

knntest_X <- knndatatest %>% select(-car_price)
knntest_Y <- knndatatest$car_price


knn_pred <- knn.reg(train = knntrain_X, test = knntest_X, y = knntrain_Y, k = 8)$pred # why the pred???

```

```{r}

#TAkes long to run 

train_control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

# Train the k-NN model using the specified cross-validation
knn_model_cv <- train(
  car_price ~ ., 
  data = knndatatrain,   # Replace with your training data
  method = "knn",        # Specify that we're using k-NN
  trControl = train_control,  # Cross-validation settings
  tuneLength = 5         # Number of different values of k to try (optional)
)

# View the results
print(knn_model_cv)
```


```{r}
# Reverse scaling for predicted prices  
predicted_pricesknn <- knn_pred * sd(all_car_adverts$car_price, na.rm = TRUE) + mean(all_car_adverts$car_price, na.rm = TRUE)

# Reverse scaling for actual prices  
actual_pricesknn <- knndatatest$car_price * sd(all_car_adverts$car_price, na.rm = TRUE) + 
                   mean(all_car_adverts$car_price, na.rm = TRUE)

# Compute MSE  
mse_knn <- mean((actual_pricesknn - predicted_pricesknn)^2)
```



Compare mse's 


```{r}


rfMSE <- resultsRF[2,2]

# Create a data frame dynamically from the MSE variables in your environment
model_names <- c("Random Forest", "SVR", "Neural Network", "k-NN", "Extreme gradiant boosting")
mse_values <- mget(c("rfMSE", "svrmse", "mse_nn", "mse_knn", "xgbmse"))

# Create the data frame
mse_data <- data.frame(Model = model_names, MSE = unlist(mse_values))

# Plot the bar plot
ggplot(mse_data, aes(x = Model, y = MSE, fill = Model)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  theme_minimal() +
  labs(title = "MSE for Different Models", y = "Mean Squared Error", x = "Model") +
  scale_fill_brewer(palette = "Set2")
```

Because SVR is so huge, lets only look at the rest

```{r}
# Define the model names, excluding "SVR"
model_names <- c("Random Forest", "Neural Network", "k-NN", "Extreme gradiant boosting")

# Extract MSE values for the models, excluding "SVR"
mse_values <- mget(c("rfMSE", "mse_nn", "mse_knn", "xgbmse"))

# Create the data frame
mse_data <- data.frame(Model = model_names, MSE = unlist(mse_values))

# Plot the bar plot
ggplot(mse_data, aes(x = Model, y = MSE, fill = Model)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  theme_minimal() +
  labs(title = "MSE for Different Models", y = "Mean Squared Error", x = "Model") +
  scale_fill_brewer(palette = "Set2")

```



Now add to a test dataframe for ease of scrutiny 

```{r}

#So some sets had 86 observations, and one had 84. So delete the two rows in other sets

final_test_setex <- final_test_set %>%
   select(-log_price, -model, -variant, -luxury_level)


removeprice <- setdiff(final_test_setex, final_test_setRF)$car_price
removemiles <- setdiff(final_test_setex, final_test_setRF)$miles

```

```{r}

#dplyr code to exludde the columns

filter(!(car_price %in% removeprice & miles %in% removemiles))

```








```{r}

comparisonframe <- final_test_setRF %>%
  mutate(Predictedxgb = predxgb,
         Predictednn = predicted_pricesNN,
         Predictedknn = predicted_pricesknn,
         PredictedRF = rf_preds_6)  # Rename appropriately


```


```{r}

#First need to format into long data

comparisonlong <- comparisonframe %>%
  pivot_longer(cols = c(PredictedRF, Predictednn, Predictedknn, Predictedxgb), 
               names_to = "Model", values_to = "Predicted_Price")

```




```{r}

ggplot(comparisonlong, aes(x = car_price, y = Predicted_Price, color = Model)) +
  geom_point(alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +  
  labs(title = "Actual vs Predicted Prices", x = "Actual Price", y = "Predicted Price") +
  scale_x_continuous(labels = scales::comma) +  
  scale_y_continuous(labels = scales::comma) + 
  theme_minimal()

```
```{r}

# Prepare the dataset
df_longluxury <- comparisonframe %>%
  select(car_price, Predictednn, luxury_level)

# Ensure Luxury_Level is numeric
df_longluxury$luxury_Level <- as.numeric(df_longluxury$luxury_level)

# Scatter plot: Actual Price (x-axis) vs Predicted Price (y-axis) with Luxury Level as color
ggplot(df_longluxury, aes(x = car_price, y = Predictednn, color = luxury_level)) +
  geom_point(size = 2, alpha = 0.7) +  # Scatter plot with transparency
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +  # Reference line y=x
  labs(
    title = "Actual vs Predicted Prices (Neural network)",
    x = "Actual Price",
    y = "Predicted Price",
    color = "Luxury Level"
  ) +
  scale_color_gradient(low = "blue", high = "red") +  
  scale_x_continuous(labels = scales::comma) +  
  scale_y_continuous(labels = scales::comma) + # Gradient from low-end to luxury
  theme_minimal()

```

Notes:

I want to use cross validation with my models as well, except for the random forest

calculate avg residual for different luxury levels 


```{r}

```

