```{r}

```{r}
# LOG PRICE VERSION - Best Subset Selection

# Step 4
target_bss_log <- "log_price"
excluded_bss_log <- c("car_price", "log_price", "model", "variant", "luxury_level")
predictors_bss_log <- setdiff(names(train_data), excluded_bss_log)

# Step 5: Generate formulas
generate_formulas <- function(p, x_vars, y_var) {
  apply(combn(x_vars, p), 2, function(vars) {
    paste0(y_var, " ~ ", paste(vars, collapse = " + "))
  })
}

predictor_range <- 3:10  # We adjusted this to balance speed and performance

# Step 6: Evaluate formulas using CV
evaluate_formula_cv <- function(formula_str, dataset, k = 5) {
  formula_obj <- as.formula(formula_str)
  folds <- sample(rep(1:k, length.out = nrow(dataset)))
  dataset <- dataset %>% mutate(folds = folds)

  mses <- numeric(k)
  r2s <- numeric(k)

  for (i in 1:k) {
    train_fold <- dataset %>% filter(folds != i)
    valid_fold <- dataset %>% filter(folds == i)

    factor_cols <- names(Filter(is.factor, train_fold))
    valid_fold <- valid_fold %>%
      mutate(across(all_of(factor_cols), ~ factor(.x, levels = levels(train_fold[[cur_column()]]))))

    model <- lm(formula_obj, data = train_fold)
    preds <- predict(model, newdata = valid_fold)
    actual <- valid_fold[[as.character(formula_obj)[2]]]

    mses[i] <- mean((actual - preds)^2)
    r2s[i] <- cor(actual, preds)^2
  }

  list(mean_mse = mean(mses), mean_r2 = mean(r2s))
}

# Step 7
bss_results_log <- data.frame(
  predictors_count = integer(),
  formula = character(),
  cv_mse = numeric(),
  cv_r2 = numeric(),
  stringsAsFactors = FALSE
)

for (p in predictor_range) {
  formulas <- generate_formulas(p, predictors_bss_log, target_bss_log)
  for (f in formulas) {
    metrics <- evaluate_formula_cv(f, train_data)
    bss_results_log <- rbind(bss_results_log, data.frame(
      predictors_count = p,
      formula = f,
      cv_mse = metrics$mean_mse,
      cv_r2 = metrics$mean_r2
    ))
  }
}

# Step 8: Select best formula
bss_results_log <- bss_results_log %>%
  mutate(score = 0.475 * (cv_mse / max(cv_mse)) +
                 0.475 * (1 - (cv_r2 / max(cv_r2))) +
                 0.05 * (predictors_count / max(predictors_count)))

best_formula_bss_log <- as.formula(bss_results_log$formula[which.min(bss_results_log$score)])

# Step 9: Fit on train+validation
train_val_data <- bind_rows(train_data, validation_data)
model_bss_log <- lm(best_formula_bss_log, data = train_val_data)

# Step 10: Test set
pred_bss_log_test <- predict(model_bss_log, newdata = test_data)
mse_bss_test_log <- mean((test_data$log_price - pred_bss_log_test)^2)
r2_bss_test_log <- cor(test_data$log_price, pred_bss_log_test)^2

# Step 11: Holy test set
pred_bss_log_holy <- predict(model_bss_log, newdata = holy_test_data)
mse_bss_holy_log <- mean((holy_test_data$log_price - pred_bss_log_holy)^2)
r2_bss_holy_log <- cor(holy_test_data$log_price, pred_bss_log_holy)^2

# CV results
mse_bss_cv_log <- min(bss_results_log$cv_mse)
r2_bss_cv_log <- bss_results_log$cv_r2[which.min(bss_results_log$score)]

```

# ACTUAL PRICE VERSION - Best Subset Selection

target_bss_actual <- "car_price"
excluded_bss_actual <- c("log_price","car_price", "model", "variant", "reg")
predictors_bss_actual <- setdiff(names(train_data), excluded_bss_actual)

bss_results_actual <- data.frame(
  predictors_count = integer(),
  formula = character(),
  cv_mse = numeric(),
  cv_r2 = numeric(),
  stringsAsFactors = FALSE
)

for (p in predictor_range) {
  formulas <- generate_formulas(p, predictors_bss_actual, target_bss_actual)
  for (f in formulas) {
    metrics <- evaluate_formula_cv(f, train_data)
    bss_results_actual <- rbind(bss_results_actual, data.frame(
      predictors_count = p,
      formula = f,
      cv_mse = metrics$mean_mse,
      cv_r2 = metrics$mean_r2
    ))
  }
}

bss_results_actual <- bss_results_actual %>%
  mutate(score = 0.475 * (cv_mse / max(cv_mse)) +
                 0.475 * (1 - (cv_r2 / max(cv_r2))) +
                 0.05 * (predictors_count / max(predictors_count)))

best_formula_bss_actual <- as.formula(bss_results_actual$formula[which.min(bss_results_actual$score)])
model_bss_actual <- lm(best_formula_bss_actual, data = train_val_data)

pred_bss_actual_test <- predict(model_bss_actual, newdata = test_data)
mse_bss_test_actual <- mean((test_data$car_price - pred_bss_actual_test)^2)
r2_bss_test_actual <- cor(test_data$car_price, pred_bss_actual_test)^2

pred_bss_actual_holy <- predict(model_bss_actual, newdata = holy_test_data)
mse_bss_holy_actual <- mean((holy_test_data$car_price - pred_bss_actual_holy)^2)
r2_bss_holy_actual <- cor(holy_test_data$car_price, pred_bss_actual_holy)^2

mse_bss_cv_actual <- min(bss_results_actual$cv_mse)
r2_bss_cv_actual <- bss_results_actual$cv_r2[which.min(bss_results_actual$score)]

```
